Neural networks have achieved remarkable success across a wide range of tasks, yet their internal representations are often entangled and difficult to modularize. While neural-symbolic reasoning and differentiable logic circuits have made progress toward combining symbolic structure with trainability~\cite{petersen2022deep, manhaeve2018deepproblog}, most architectures still struggle to express logic-based computation in a compositional and scalable way.

In this work, we introduce \textit{Operand Selective Logic Gated Networks (OSLGN)}, a symbolic neural architecture where each layer is a collection of binary logic gates. Each gate selects two operands from the previous layer via a differentiable $\arg\max$ mechanism (using a straight-through estimator), and applies one of 16 predefined binary logic operations (e.g., AND, OR, XOR). While prior works have focused on differentiable operator selection~\cite{petersen2022deep}, our key contribution is to enable operand-level routing. This allows each gate to learn not only what operation to apply, but also which inputs to apply it to—forming discrete, modular logic circuits.

To encourage the emergence of localized symbolic structure, we initialize operand selectors with a Gaussian proximity bias inspired by small-world connectivity~\cite{watts1998collective, javaheripi2019swnet}. This inductive prior promotes compositional logic among nearby nodes while still allowing long-range dependencies to emerge through learning.

We further interpret our architecture as a fine-grained form of Mixture-of-Experts~\cite{shazeer2017outrageously}, where selection happens at the operand level within each logic gate. This facilitates sparse symbolic computation with minimal overhead.

\paragraph{Contributions.}
\begin{itemize}
    \item We propose Operand-Selective Logic Gated Networks (OSLGN), a symbolic neural architecture that builds differentiable logic circuits via operand and operator selection.
    
    \item We introduce an operand selection mechanism that enables gate-level modularity. While our operator routing builds on prior differentiable logic gates~\cite{petersen2022deep}, operand selection is independently learned and crucial for symbolic structure.
    
    \item We incorporate a proximity-biased operand initialization scheme inspired by small-world networks, promoting symbolic locality and stable training dynamics.
    
    \item We show that OSLGN can be fully translated into compact logic expressions post-training, demonstrating the feasibility of distilling a neural network into an explicit symbolic circuit.
\end{itemize}
